Adapting Duplicate Filters to a Sliding Window Context

### 2020-03-18 qhtv2.tex
Clarifications 3.2

\begin{theorem}
	\label{thm:queue-pfp}
	 Let $\FP_{\mathcal Q, m}^w$ be the probability of false positive (def. \ref{def:pfp-pfn}) of $\mathcal Q$ after $m$ insertions, over a sliding window of size $w = cL$, we have:
	\begin{equation}
		\label{eq:queue-pfp}
		\FP_{\mathcal Q, m}^w = 1 - \left(1 - \FP_{\mathcal F,c}\right)^{L-1} \left(1 - \FP_{\mathcal F, m \bmod c }\right)
	\end{equation}
	where $\bmod$ is the modulo operator and 
	 $\FP_{\mathcal F,m}$ be the probability of false positive of any of $\mathcal Q$'s subfilters $\mathcal F_i$ after $m$ insertions.
\end{theorem}

@@

\subsubsection{Probability of False Negative}\label{sub:fnr}
\begin{theorem}
	Let $\mathcal Q$ be a queueing filter, consisting of $L$ underlying subfilters $\mathcal F$, each filled with up to $c$ elements. Let $\FN_{\mathcal F,m}$ be the probability of false negative of the subfilter $\mathcal F$ after $m$ insertions, let $\FN_{\mathcal Q, m}^w$ be the probability of false negative of $\mathcal Q$ after $m$ insertions on a sliding window of size $w$.

	For $p_m = \frac{1-\left(1-\frac 1{|\Gamma|}\right)^m}{1-\left(1-\frac 1{|\Gamma|}\right)^w}$, we have 
	\begin{align*}
	\FN_{\mathcal Q, m}^w =& \left[p_c\FN_{\mathcal F,c} + \left(1-p_c\right)\left(1-\FP_{\mathcal F,c}\right)\right]^{L-1}
	\\&\qquad
	 \cdot \left[p_{m\%c}\FN_{\mathcal F,m\%c} + \left(1-p_{m\%c}\right)\left(1-\FP_{\mathcal F,m\%c}\right)\right]
	\end{align*}

	where $\%$ is the modulo operator.
\end{theorem}

\begin{proof}
	Let $E = \{e_1, \dotsc, e_m, \dotsc\}$ be a stream, let $w$ be a sliding window and let $e^\star \wdup E$.
	We first observe that if $cL > w$, there might be some duplicates of $e^\star$ that happened before the current sliding window, but might still be remembered by the oldest filter. However, as argued in the previous section, the probability of such an event is negligible.

	Hence, we consider that the only duplicates are part of the sliding window. $e^\star$ is a false negative if and only if all subfilters $\mathcal F_i$ answer $\mathcal F_i.\mathsf{Detect}(e^\star) = \UNSEEN$.

	There can be two cases:
	\begin{itemize}
		\item $e^\star$ is present in $\mathcal F_i$'s sub-sliding window
		\item $e^\star$ is not present in $\mathcal F_i$'s sub-sliding window
	\end{itemize}

In the first case, $\mathcal F_i.\mathsf{Detect}(e^\star)$ returns \UNSEEN if and only if $e^\star$ is a false negative for $\mathcal F_i$. Hence, in this case, the probability of $\mathcal F_i$ to return \UNSEEN is $\FN_{\mathcal F,c}$. For the subfilter $\mathcal F_0$, however, the probability is $\FN_{\mathcal F, m\%c}$.

In the second case, $\mathcal F_i.\mathsf{Detect}(e^\star)$ returns \UNSEEN if and only if $e^\star$ is not a false positive for $\mathcal F_i$. Hence, in this case, the probability of $\mathcal F_i$ to return \UNSEEN is $1-\FP_{\mathcal F,c}$. For the subfilter $\mathcal F_0$, however, the probability is $1-\FP_{\mathcal F, m\%c}$.

Now, the probability $p_c$ for $e^\star$ to be in subfilter $\mathcal F_i$ (which has had $c$ insertions) sub-sliding window is

\begin{align*}
	p_c =& Pr[\text{$e^\star$ is in }\mathcal F_i \text{ sub-sliding window | } e^\star \wdup E]\\
	=& \frac{Pr[\text{$e^\star$ is in }\mathcal F_i \text{ sub-sliding window } \cap e^\star \wdup E]}{Pr[e^\star \wdup E]}\\
	=& \frac{Pr[\text{$e^\star$ is in }\mathcal F_i \text{ sub-sliding window}]}{Pr[e^\star \wdup E]}\\
	=& \frac{1 - Pr[\text{$e^\star$ is not in }\mathcal F_i \text{ sub-sliding window }]}{1 - Pr[e^\star \notwdup E]}\\
	p_c=& \frac{1 - \left(1 - \frac 1{|\Gamma|}\right)^c}{1 - \left(1-\frac{1}{|\Gamma|}\right)^w}
\end{align*}

Finally, we get that the probability of a false positive for $\mathcal Q$ over the sliding window $w$ is 
\begin{align*}
\FN^w_{\mathcal Q, m} =& \left[p_c\FN_{\mathcal F,c} + (1-p_c)(1-\FP_{\mathcal F,c})\right]^{L-1}
\\&\quad
\cdot \left[p_{m\%c}\FN_{\mathcal F,m\%c} + (1-p_{m\%c})(1-\FP_{\mathcal F,m\%c})\right]\\
\end{align*}
\end{proof}

This probability is quite complicated and does not allow, in the general case, for an in-depth analysis of the probability of false negative.

>>>>

\begin{theorem}
	\label{thm:queue-pfp}
	 Let $\FP_{\mathcal Q, m}^w$ be the probability of false positive (def. \ref{def:pfp-pfn}) of $\mathcal Q$ after $m > w$ insertions, over a sliding window of size $w = cL$, we have:
	\begin{equation}
		\label{eq:queue-pfp}
		\FP_{\mathcal Q, m}^w = 1 - \left(1 - \FP_{\mathcal F,c}\right)^{L-1} \left(1 - \FP_{\mathcal F, m \bmod c }\right)
	\end{equation}
	where $\bmod$ is the modulo operator and 
	 $\FP_{\mathcal F,m}$ be the probability of false positive of any of $\mathcal Q$'s subfilters $\mathcal F_i$ after $m$ insertions.
\end{theorem}

@@

\subsubsection{Probability of False Negative}\label{sub:fnr}
\begin{theorem}
	\label{thm:queue-pfn}
	Let $\FN_{\mathcal Q, m}^w$ be the probability of false negative of $\mathcal Q$ after $m > w$ insertions on a sliding window of size $w = cL$, we have
	\begin{align*}
	\FN_{\mathcal Q, m}^w 
	= & \left[p_c\FN_{\mathcal F,c} + \left(1-p_c\right)\left(1-\FP_{\mathcal F,c}\right)\right]^{L-1} \\
	 \times & \left[p_{m\bmod c}\FN_{\mathcal F,m\bmod c} + \left(1-p_{m\bmod c}\right)\left(1-\FP_{\mathcal F,m\bmod c}\right)\right]
	\end{align*}
	where $\bmod$ is the modulo operator, $\FN_{\mathcal F,m}$ is the probability of false negative of any subfilter $\mathcal F$ after $m$ insertions, and
	\begin{equation*}
		p_m = \frac{1-\left(1-\frac 1{|\Gamma|}\right)^m}{1-\left(1-\frac 1{|\Gamma|}\right)^w}.
	\end{equation*}
\end{theorem}

\begin{proof}
	Let $E = \{e_1, \dotsc, e_m, \dotsc\}$ be a stream, let $w$ be a sliding window and let $e^\star \wdup E$.

	Then $e^\star$ is a false negative if and only if all subfilters $\mathcal F_i$ answer $\mathcal F_i.\mathsf{Detect}(e^\star) = \UNSEEN$. There can be two cases:
	\begin{itemize}
		\item $e^\star$ is present in $\mathcal F_i$'s sub-sliding window;
		\item $e^\star$ is not present in $\mathcal F_i$'s sub-sliding window.
	\end{itemize}
In the first case, $\mathcal F_i.\mathsf{Detect}(e^\star)$ returns \UNSEEN if and only if $e^\star$ is a false negative for $\mathcal F_i$. This happens with probability $\FN_{\mathcal F,c}$ by definition, except for $\mathcal F_0$, for which the probability is $\FN_{\mathcal F, m\bmod c}$.

In the second case, $\mathcal F_i.\mathsf{Detect}(e^\star)$ returns \UNSEEN if and only if $e^\star$ is not a false positive for $\mathcal F_i$, which happens with probability $1-\FP_{\mathcal F,c}$, execpt for $\mathcal F_0$, for which the probability is $1-\FP_{\mathcal F, m\bmod c}$.

Finally, each event is weighted by the probability $p_c$ that $e^\star$ is in $\mathcal F_i$'s sub-sliding window:
\begin{align*}
	p_c 
	& = \Pr[\text{$e^\star$ is in }\mathcal F_i \text{ sub-sliding window | } e^\star \wdup E]\\
	& = \frac{\Pr[\text{$e^\star$ is in }\mathcal F_i \text{ sub-sliding window } \cap e^\star \wdup E]}{\Pr[e^\star \wdup E]}\\
	& = \frac{\Pr[\text{$e^\star$ is in }\mathcal F_i \text{ sub-sliding window}]}{\Pr[e^\star \wdup E]}\\
	& = \frac{1 - \Pr[\text{$e^\star$ is not in }\mathcal F_i \text{ sub-sliding window }]}{1 - \Pr[e^\star \notwdup E]}\\
	& = \frac{1 - \left(1 - \frac 1{|\Gamma|}\right)^c}{1 - \left(1-\frac{1}{|\Gamma|}\right)^w}
\end{align*}
This concludes the proof.
\end{proof}
\paragraph{Remark.} As previously, the effect of $w < cL$ is negiglible for all practical purposes and Theorem~\ref{thm:queue-pfn} is considered a good approximation in that regime.

\subsection{FNR and FPR}

\todo[TODO]

\section{Optimising queuing filters}
